Abstract

### 2025-10-24 output/Response_to_action_editor_reviewers-2.Rmd

The only suggestions not accepted are:

In p. 4 (line 2), we think that "the two dimensions [...] _combine_ in the second group" does not reflect accurately the fact that the two dimensions are collapsed into a single one when the correlation is 1.
We have therefore decided to just drop this clause, as it is redundant with the previous one which states that "the two dimensions are _perfectly aligned_".
This sentence now reads:

> [...] because the two dimensions are *perfectly aligned* in the second group,
the item `r MDISC_SYM` should be the algebraic sum of both components, [...]

In p. 12 (line 8), simply dropping "already" made the sentence lose the intended
meaning that the implicit assumption of orthogonality in @reckase_difficulty_1985 is already present in Equation 3, and not just from Equation 7 (p. 404) on, as one might think in the first place.
We deem this point important because it took us quite some effort to notice that generalizing the transformation to polar coordinates to non-orthonormal bases implied reformulating this equation.
In an attempt to keep the intended meaning while improving the sentence structure, we have restated it as:

> [...] this shows that the orthogonality of the latent space in @reckase_difficulty_1985 is assumed as early on as in Equation 3 (p. 403) [...]

### 2025-10-24 output/Main_text.Rmd
Update explanation of axis lableing in Figure 1
Implements the suggestion by @mluzardov after discussion.

Up to now, all our derivations have considered $`r N_DIMS`$-dimensional spaces;
that is, multidimensional parameters and item vector coordinates
can be computed in an arbitrary large number of dimensions.
However, as plotting more than two dimensions
is difficult on a bidimensional display
(and hardly possible at all for more than three dimensions with the
currently available technology),
we limit ourselves to the bidimensional case here.
The example in Figure \@ref(fig:item-plot-out) showcases a representation of
a set of items with two different latent trait distributions.
(The axes are labeled as $`r TRAIT_COMP_1`$ and $`r TRAIT_COMP_2`$
for simplicity, but note that they represent the coordinates of
a latent trait point expressed in test space coordinates,
that is, $`r TRAIT_COORD_1`$ and  $`r TRAIT_COORD_2`$.)
Their parameters are shown in Table \@ref(tab:example-items-table-out),
both in the original `r MODEL_ACRONYM` metric and
in the multidimensional metric under the two cases,
which differ in the latent trait correlation.
In the first case (columns $`r CORR_ORTH_OUT`$),
the two latent dimensions are independent;
in the second one ($`r CORR_OBL_OUT`$),
the correlation between them is .5.
The same `r MODEL_ACRONYM` parameters are used in both cases,
but their (correlation-based) multidimensional parameters change
with the correlation.
If we used the agnostic version instead, the multidimensional parameters,
given by the values under $`r CORR_ORTH_OUT`$, would be the same in both cases.

>>>>

Up to now, all our derivations have considered $`r N_DIMS`$-dimensional spaces;
that is, multidimensional parameters and item vector coordinates
can be computed in an arbitrary large number of dimensions.
However, as plotting more than two dimensions
is difficult on a bidimensional display
(and hardly possible at all for more than three dimensions with the
currently available technology),
we limit ourselves to the bidimensional case here.
The example in Figure \@ref(fig:item-plot-out) showcases
the test space representation of a set of items
under two different latent trait distributions.
In formal terms, the proper axis labeling of this figure
should correspond to the vectors that make up basis $`r TEST_SPACE_BASIS`$,
in this case, $`r ITEM_BASIS_VECTOR_FIRST`$ and $`r ITEM_BASIS_VECTOR_2`$;
the coordinates represent the coefficients that multiply each basis vector to
obtain the corresponding coordinate in the test space.
However, these coefficients correspond to the values of
a vector in the test space, be it $`r DISCR_COORDS`$ for $`r DISCR_VECTOR`$,
or $`r TRAIT_COORDS_TRANSF_TS`$ for $`r TRAIT_VECTOR`$
(as given by Equation \@ref(eq:trait-coords-spaces-transform)).
For the sake of simplicity,
we label the axes as $`r TRAIT_COMP_1`$ and $`r TRAIT_COMP_2`$,
highlighting the comparison of the items with the latent trait vectors.
However, note that these coordinates actually correspond to
the components of $`r TRAIT_COORDS_TRANSF_TS`$, i.e.,
$`r TRAIT_COORD_1`$ and  $`r TRAIT_COORD_2`$.

The parameters of the items in Figure \@ref(fig:item-plot-out) are shown
in Table \@ref(tab:example-items-table-out),
both in the original `r MODEL_ACRONYM` metric and
in the multidimensional metric under the two cases,
which differ in the latent trait correlation.
In the first case (columns $`r CORR_ORTH_OUT`$),
the two latent dimensions are independent;
in the second one ($`r CORR_OBL_OUT`$),
the correlation between them is .5.
The same `r MODEL_ACRONYM` parameters are used in both cases,
but their (correlation-based) multidimensional parameters change
with the correlation.
(If we used the agnostic version instead, the multidimensional parameters,
given by the values under $`r CORR_ORTH_OUT`$, would be the same in both cases.)
